---
title: "Power to the People: Data Transparency in Crytocurrency"
subtitle: "Developing the 'Ignition Trading Bot'"
author: "Joshua Allessio"
date: "2023-10-21"
output: html_document
---

```{r include=FALSE}
# Load Packages
library(tidyr)
library(dplyr)
library(here)
library(readr)
rm(list = ls()) # Clear environment so data doesn't get mixed up
```


### Import the Dataset
I collected this data set over the course of a few weeks using the Coinalyze.net free API. Every time I saw a crypto token pump, I saved a .csv of the last day or so worth of it's price and Open Interest history at 1-minute granularity. Open Interest is a measure of market participation and represents the total number of open futures contracts. Rising open interest is a strong bullish sign. Rapidly rising Open Interest can be a very strong sign. We can use this data to observe and make conclusions about how crypto coin pumps happen. I also saved plenty of instances where the crypto token did NOT increase greatly, and instances when the chart *almost* looked like the charts that I do want to trade. This way the resultant strategy from this backtest is resilient to false-positives.
```{r}
filenames <- list.files(path = "C:/Users/jalle/Documents/AAA School Stuff/Fall2023/Crypto_Screenr/Data/1m") 
Crypto_Data <- list()    #Initialize an empty list to later organize all of our data
for (i in filenames) {
  Crypto_Data[[i]] = data.frame(read_csv(show_col_types = FALSE, here("C:/Users/jalle/Documents/AAA School Stuff/Fall2023/Crypto_Screenr/Data/1m",i)))
}
```


There are several questions we need to answer here to paint the picture we are trying to see. Overall, the main question is, "Does a significant increase, in either a sustained or rapid manner, in Open Interest - strongly correlate to a great increase in price as well? If so, what kind of threshold should we use to look for OI increases that are most likely to lead to a price increase?"


### Making Observations
The question to answer here is, "When OI increases by X% or more in just 15 minutes, what tends to happen to price afterwords?" Clarifying sub-questions we have for this question are, "What was the maximum price increase from the signal? What was the maximum price decrease from the signal? What was the maximum drawdown in between the signal and the maximum increase?"

1. Get the "current" OI open. We will begin at the 15th observation and look backwards, as to simulate a real life environment where we can only look backwards, not forwards.
2. Calculate the percent change to the "current" OI value from the previous one, and from the previous 15 observations. We can store these values in their own array data frame. 
3. Determine if any of those values are greater than X% in X minutes. We can start with at least 8% in 15 minutes.
4. If the above condition is satisfied, exit that calculation loop. This is where we would "enter" a position.
5. For now, let's just see what was the maximum percent price increase (and decrease) from that point. 
```{r}
trades = data.frame()           # Initialize an empty data frame that we will need later
for (i in Crypto_Data) {        # In this loop, i becomes each one of our data frames that we collected earlier
  for (j in 15:length(i$OI_open)) {   # In this loop, j is a counter variable representing the current row number
    OI_now <- i$OI_open[j]                # This represents the "current" OI value
    OI_recent <- i$OI_open[(j-15):(j-1)]  # This represents the 15 most recent OI values
    delta <- 100*(OI_now-OI_recent)/OI_recent # This is the %change in OI from each of the last 15 values
    if (any(delta > 6)) {                     # Determine if OI has risen 8% in the last 15 minutes
    RowNum <- j                               # Note the row where the condition became satisfied
    break                                     # Exit the loop. This is where we would enter a trade.
    }
  }
                                              # Here, we we make some simple observations about where price goes after a rapid OI increase.
  if (RowNum < length(i$OI_open)) {           # We are now back in the outer loop. i is the chart we are trading and j is the entry point.
    currentval <- i$price_close[j]
    # print(RowNum)
    maxval <- max(i$price_close[j:length(i$price_close)])
    minval <- min(i$price_close[j:match(maxval, i$price_close)])  
    MaxGain = 100*(maxval - currentval)/currentval
    MaxLoss = 100*(minval - currentval)/currentval
    
    # Create a data frame containing all the trades that would have been taken
    if (!is.na(MaxGain) && !is.na(MaxLoss) && MaxGain != 0 && MaxLoss != 0) {
      # Calculate the time elapsed for the "trade"
      t1 <- i$timestamp[j]                            |> as.POSIXct(origin = "1970-01-01")
      t2 <- i$timestamp[match(maxval, i$price_close)] |> as.POSIXct(origin = "1970-01-01")
      ElapsedHours <- as.numeric(difftime(t2, t1, units = "hours")) |> round(1)
      trade <- data.frame(MaxGain, MaxLoss, ElapsedHours)
      trades <- rbind(trades, trade)
    } else {
    }
  }
}
trades <- trades |> mutate(RiskRewardRatio = MaxGain/abs(MaxLoss)) # Calculate the Hypothetical R for each trade
print(trades)
```

### Optimization
The backtest is looking great so far. But why not take it one step further? The code block below parameterizes the above code block and uses an optimization function to attempt to discover better parameters. I chose to optimize for the *maximum median risk:reward ratio*. I opted not to optimize the mean risk:reward ratio because the mean is more susceptible to outliers. Optimizing the median incentivizes the algorithm to maximize the risk:reward ratio for EVERY trade, and disincentivizes finding a condition with one outstanding trade and a dozen meaningless ones.
```{r}
optimize <- function (params) {
local_lookback <- params[1]     # Parameterize the local lookback duration
OI_threshold <- params[2]       # Parameterize the OI increase threshold
trades = data.frame()           # Initialize an empty data frame that we will need later
for (i in Crypto_Data) {        # In this loop, i becomes each one of our data frames that we collected earlier
  for (j in local_lookback:length(i$OI_open)) {   # In this loop, j is a counter variable representing the current row number
    OI_now <- i$OI_open[j]                # This represents the "current" OI value
    OI_recent <- i$OI_open[(j-local_lookback):(j-1)]  # This represents the n most recent OI values
    delta <- 100*(OI_now-OI_recent)/OI_recent # This is the %change in OI from each of the last n values
    if (any(delta > OI_threshold)) {         # Determine if OI has risen 8% in the last n minutes
    RowNum <- j                               # Note the row where the condition became satisfied
    break                                     # Exit the loop. This is where we would enter a trade.
    }
  }
                                              # Here, we we make some simple observations about where price goes after a rapid OI increase.
  if (RowNum < length(i$OI_open)) {           # We are now back in the outer loop. i is the chart we are trading and j is the entry point.
    currentval <- i$price_close[j]
    # print(RowNum)
    maxval <- max(i$price_close[j:length(i$price_close)])
    minval <- min(i$price_close[j:match(maxval, i$price_close)])  
    MaxGain = 100*(maxval - currentval)/currentval
    MaxLoss = 100*(minval - currentval)/currentval
    
    # Create a data frame containing all the trades that would have been taken
    if (!is.na(MaxGain) && !is.na(MaxLoss) && MaxGain != 0 && MaxLoss != 0) {
      # Calculate the time elapsed for the "trade"
      t1 <- i$timestamp[j]                            |> as.POSIXct(origin = "1970-01-01")
      t2 <- i$timestamp[match(maxval, i$price_close)] |> as.POSIXct(origin = "1970-01-01")
      ElapsedHours <- as.numeric(difftime(t2, t1, units = "hours")) |> round(1)
      trade <- data.frame(MaxGain, MaxLoss, ElapsedHours)
      trades <- rbind(trades, trade)
    } else {
    }
  }
}
trades <- trades |> mutate(RiskRewardRatio = MaxGain/abs(MaxLoss)) # Calculate the Hypothetical R for each trade
return(-median(trades$MaxGain))  #since optim() will seek to minimize, we will seek to minimize the negtive median in order to maximize.
}
# Set up our optimization parameters
lower_bounds <- c(9, 4)    # Lower bounds for each parameter
upper_bounds <- c(25, 15)  # Upper bounds for each parameter
guess = c(1, 1)            # Initial guess for the function to begin
optimal_params <- optim(par = guess, fn = optimize, method = "L-BFGS-B", lower = lower_bounds, upper = upper_bounds)
print(paste("The optimal local lookback period is", optimal_params$par[1], "minutes, and the optimal threshold for the OI increase is", optimal_params$par[2], "percent"))

```
By iterating with a few different initial guesses, we find a few optimal pairs of local_lookback and OI_threshold. 15 and 6 is one, but we also discover 9 and 4. Meaning - searching the past 9 minutes for a 4% OI increase or greater may be an optimal strategy. This is great news, because a shorter lookback time AND a smaller OI increase threshold means that we will be entering trades sooner and getting a better entry price. But wait, there's more! If you go back to the previous code block and input these new parameters, we notice two things about the new "trades" data frame.
```{r echo = FALSE}
trades = data.frame()           # Initialize an empty data frame that we will need later
for (i in Crypto_Data) {        # In this loop, i becomes each one of our data frames that we collected earlier
  for (j in 9:length(i$OI_open)) {   # In this loop, j is a counter variable representing the current row number
    OI_now <- i$OI_open[j]                # This represents the "current" OI value
    OI_recent <- i$OI_open[(j-9):(j-1)]  # This represents the 15 most recent OI values
    delta <- 100*(OI_now-OI_recent)/OI_recent # This is the %change in OI from each of the last 15 values
    if (any(delta > 4)) {                     # Determine if OI has risen 8% in the last 15 minutes
    RowNum <- j                               # Note the row where the condition became satisfied
    break                                     # Exit the loop. This is where we would enter a trade.
    }
  }
                                              # Here, we we make some simple observations about where price goes after a rapid OI increase.
  if (RowNum < length(i$OI_open)) {           # We are now back in the outer loop. i is the chart we are trading and j is the entry point.
    currentval <- i$price_close[j]
    # print(RowNum)
    maxval <- max(i$price_close[j:length(i$price_close)])
    minval <- min(i$price_close[j:match(maxval, i$price_close)])  
    MaxGain = 100*(maxval - currentval)/currentval
    MaxLoss = 100*(minval - currentval)/currentval
    
    # Create a data frame containing all the trades that would have been taken
    if (!is.na(MaxGain) && !is.na(MaxLoss) && MaxGain != 0 && MaxLoss != 0) {
      # Calculate the time elapsed for the "trade"
      t1 <- i$timestamp[j]                            |> as.POSIXct(origin = "1970-01-01")
      t2 <- i$timestamp[match(maxval, i$price_close)] |> as.POSIXct(origin = "1970-01-01")
      ElapsedHours <- as.numeric(difftime(t2, t1, units = "hours")) |> round(1)
      trade <- data.frame(MaxGain, MaxLoss, ElapsedHours)
      trades <- rbind(trades, trade)
    } else {
    }
  }
}
trades <- trades |> mutate(RiskRewardRatio = MaxGain/abs(MaxLoss)) # Calculate the Hypothetical R for each trade
print(trades)
```

One: we end up taking way more trades. This is actually a good thing. When we have a positive expected value, taking more trades has the effect of smoothing out the variance of our equity curve. So the trading account should experience less volatility and drawdown. 

Two: We have successfully increased the median R:R per trade; with one trade reaching a theoretical maximum R:R of 97! (Not a realistic expectation, but fun to think about)

Overall, this optimization block was very useful and gives us much less stringent constraints to work around in the Risk Management section below. 



### Risk Management
It looks like we are onto something so far, but the project is far from complete. So far, we have discovered one of the three components of a trade. The three components are: Entry, Take-profit(s), and Stop-loss. 

Ideally we want to exit losing trades quickly and keep losses small, and ideally we want to let winning trades run. Exiting losing trades is done with a "Stop Loss." A stop loss is a risk management strategy where we decide on a price at which we no longer believe the trade idea is correct, and we place an exit order at that price. 

Ideally, the stop-loss is volatility-weighted. When I started collecting data for this project, I opted to only keep the 1-minute price opens and closes - instead of open, close, high, and low. OHLC would be more useful for measuring volatility than just O & C. I had never done a data science project before and I thought I might want opt to store less data and perform less calculations. Little did I know just how miniscule my project was on the scale of my computer's capabilities - I should have just kept all the data and then some. 

Regardless, I am pleased to observe that drawdown is typically very low, and the base average R (risk-reward ratio) is very high. This gives me a lot of extra room to build in risk management and still stay profitable. Thankfully, because of the nature of this strategy, the theoretical entry points are exactly at the beginning of violent upwards movements. You can see this in the fact that most of the drawdown values are very low. So, I am comfortable setting a hard stop-loss at -n% or so instead of trying to figure out how to weight it by volatility. We will do some more analysis to figure out the value of n. Maybe we can even have a tiered system where there are a few options for n based on the agnitude of the local OI increase.

The other consideration is the take-profit order(s). The model above finds the absolute maximum gain from the entry. We need to model a reasonable exit, instead of using the absolute maximum. Now, we will create some visualizations to get to a starting point for our stop-loss and take-profit strategies. 

```{r}



```

